---
title: "Group Project Mobile Price Classification"
author: "Group A14"
date: "2022-12-10"
output:
  pdf_document: 
    toc: yes
    toc_depth: 4
  html_document: 
    highlight: haddock
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
---
Notes: 
1.cleaned data: baked_train1, baked_test1
2.set.seed(14)
3.to tune hyperparameter, use cv accuracy，cv fold = 10.


```{r}
pacman::p_load(ggplot2,dplyr,caret,rsample,nnet,MASS,car,recipes,readr,class,DMwR2)
```

```{r eval=FALSE, include=FALSE}
#data <- read.csv("train.csv")
```

```{r}
data <- read.table('A14_Project_Data.txt', skip = 1, header =TRUE, sep ='\t')
head(data)
```


```{r}
str(data)
dim(data)
```

# check missing value
```{r}
sum(is.na(data))
```

# change multiple price_range to binary price_range
```{r}
data$price <- as.integer(data$price_range>1)
```

```{r}
data %>%
  group_by(price_range,price) %>%
  summarise(number = n(),.groups="drop") %>%
  arrange(price_range)
```
# make a new data without price_range
```{r}
data1 <- data[,-21]
str(data1)
```

# heat-map
```{r}
ggplot(data = data %>% select_if(is.numeric) %>% cor() %>% reshape2::melt(),
        aes(x = Var1 ,y = Var2, fill = value)) +
  geom_tile(color="white",size=0.1) +
  xlab("") +
  ylab("") +
  guides(fill = guide_legend(data = "Correlation")) +
  scale_fill_gradient( low = "#56B1F7", high = "#132B43") +     #lightblue to darkblue
  theme(axis.text.x = element_text(angle = 25, hjust = 1))
```

```{r}
#raw vs price_range
data$price <- as.factor(data$price)
ggplot(data=data, aes(x=price, y=ram)) + 
geom_boxplot()
```

```{r}
#battery_power vs price_range
ggplot(data=data, aes(x=price, y=battery_power)) + 
geom_boxplot()
```
## impute missing value : sc_w= 0
```{r}
data1[data1$sc_w==0,]$sc_w=NA
data1=knnImputation(data1)
```

```{r}
summary(data1)
```
#check if sc_w = 0 is imputed with other values
#data1[13:14,]


# stratified sampling 
```{r}
set.seed(14)
split <- initial_split(data1,prop=0.8,strata = "price")
train <- training(split)
test <- testing(split)
```

# feature engineering
as the data has already done the 0-1 variable, we should take care when do feature engineering
```{r}
blueprint <- recipe(price~.,data=train[,-c(2,4,6,18,19,20)]) %>% 
  # don't change binary variables c(2,4,6,18,19,20).
  step_nzv(all_nominal()) %>%
  step_scale(all_numeric(),-all_outcomes()) %>%
  step_center(all_numeric(),-all_outcomes()) %>%
  step_dummy(all_nominal(),-all_outcomes(),one_hot = TRUE)
```

```{r}
prepare <- prep(blueprint,training =train[,-c(2,4,6,18,19,20)])
```

```{r}
baked_train <- bake(prepare,new_data = train[,-c(2,4,6,18,19,20)])
baked_test <- bake(prepare,new_data = test[,-c(2,4,6,18,19,20)])
```


```{r}
# combine baked data with original binary variables c(2,4,6,18,19,20)
baked_train1 <- cbind(baked_train,train[,c(2,4,6,18,19,20)])
baked_test1 <- cbind(baked_test,test[,c(2,4,6,18,19,20)])
```

```{r}
#change all binary variables to factor for baked_train 
#baked_train1[,15:21]
baked_train1$price <- as.factor(baked_train1$price)
baked_train1$blue <- as.factor(baked_train1$blue)
baked_train1$dual_sim <- as.factor(baked_train1$dual_sim)
baked_train1$four_g <- as.factor(baked_train1$four_g)
baked_train1$three_g <- as.factor(baked_train1$three_g)
baked_train1$touch_screen <- as.factor(baked_train1$touch_screen)
baked_train1$wifi <- as.factor(baked_train1$wifi)
head(baked_train1)
```

```{r}
#change all binary variables to factor for test_train 
baked_test1$price <- as.factor(baked_test1$price)
baked_test1$blue <- as.factor(baked_test1$blue)
baked_test1$dual_sim <- as.factor(baked_test1$dual_sim)
baked_test1$four_g <- as.factor(baked_test1$four_g)
baked_test1$three_g <- as.factor(baked_test1$three_g)
baked_test1$touch_screen <- as.factor(baked_test1$touch_screen)
baked_test1$wifi <- as.factor(baked_test1$wifi)
head(baked_test1)
```

# logistic regression
```{r warning = FALSE}
fit_logit = glm(price~.,family = binomial(logit), baked_train1)
summary(fit_logit)
```
```{r}
p1_logit= vip::vip(fit_logit, num_feature = 20, scale = TRUE)
gridExtra::grid.arrange(p1_logit, nrow = 1)
```

```{r warning = FALSE}
library(car)
Anova(fit_logit)
```
battery_power, mobile_wt, pc, px_height, px_width, ram are 6 siginificant variables.
optimal model: price~battery_power+mobile_wt+pc+px_height+px_width+ram

```{r}
## pdp
library(pdp)
library(ggplot2)

pdp_pred <- function(object, newdata){
  pred_Obj <- ranger::predictions(predict(object, newdata))
  results <- mean(as.vector(pred_Obj))
  return(results)
}

pd_values1 <- partial(
  fit_logit,
  train = baked_train1,
  pred.var = "ram",
  grid.resolution = 20
)

pd_values2 <- partial(
  fit_logit,
  train = baked_train1,
  pred.var = "battery_power",
  grid.resolution = 20
)

autoplot(pd_values1, rug = TRUE, train = baked_train1)
autoplot(pd_values2, rug = TRUE, train = baked_train1)
```


```{r warning = FALSE}
fit_logit_final = glm(price~battery_power+mobile_wt+pc+px_height+px_width+ram,family = binomial(logit), baked_train1)
summary(fit_logit_final)
```

```{r warning = FALSE}
# do the cross validation
#install.packages('boot')
library(boot)
set.seed(14)
fit_logit_cv <- cv.glm(baked_train1,fit_logit_final, K=10) 
fit_logit_cv$delta[1]
```

```{r}
# the cv accuracy
fit_logit_cv_accuracy <- 1- 0.005481233
fit_logit_cv_accuracy
```
the accuracy in train data is 99.45%.

```{r}
# calculate the accuracy in train data
pred_train1<-predict(fit_logit_final,newdata = baked_train1, type = "response")
bio_train1<-ifelse(pred_train1>0.5,1,0)
table(baked_train1$price,bio_train1)
```
```{r}
accuracy_train1<-1-sum(bio_train1!=baked_train1$price)/1600
accuracy_train1
```
the accuracy in train data is 99.44%.
```{r}
# calculate the accuracy in test data
pred_test1<-predict(fit_logit_final,newdata = baked_test1, type = "response")
bio_test1<-ifelse(pred_test1>0.5,1,0)
table(baked_test1$price,bio_test1)
```
```{r}
accuracy_test1<-1-sum(bio_test1!=baked_test1$price)/400
accuracy_test1
```
the accuracy in test data is 98.25%.

```{r}
# final model result
performance.log <- data.frame(
  method = 'Logistic',
  train_accuracy = accuracy_train1,
  test_accuracy = accuracy_test1
  )
performance.log
```

# poly logistic regression

```{r warning = FALSE}
# we choose two features which have the biggest impact on the price to do the regression
fit_poly1 <- glm(price ~ poly(battery_power,3) +poly(ram,3), data = baked_train1, family = binomial)
summary(fit_poly1)
```

```{r}
p1_poly= vip::vip(fit_poly1, num_feature = 20, scale = TRUE)
gridExtra::grid.arrange(p1_poly, nrow = 1)
```
The final model is price ~ battery_power+ram+ram^2

```{r}
fit_poly_final = glm(price~battery_power+poly(ram,2),family = binomial(logit), baked_train1)
summary(fit_poly_final)
```

```{r warning = FALSE}
# do the cross validation
library(boot)
set.seed(14)
fit_poly_cv <- cv.glm(baked_train1,fit_poly_final, K=10) 
fit_poly_cv$delta[1]
```
```{r}
# the cv accuracy
fit_poly_cv_accuracy <- 1- 0.04526164
fit_poly_cv_accuracy
```
the accuracy in train data is 95.47%.

```{r}
# calculate the accuracy in train data
pred_train2<-predict(fit_poly_final,newdata = baked_train1, type = "response")
bio_train2<-ifelse(pred_train2>0.5,1,0)
table(baked_train1$price,bio_train2)
```
```{r}
accuracy_train2<-1-sum(bio_train2!=baked_train1$price)/1600
accuracy_train2
```
the accuracy in train data is 93.44%.

```{r}
# calculate the accuracy in test data
pred_test2<-predict(fit_poly_final,newdata = baked_test1, type = "response")
bio_test2<-ifelse(pred_test2>0.5,1,0)
table(baked_test1$price,bio_test2)
```
```{r}
accuracy_test2<-1-sum(bio_test2!=baked_test1$price)/400
accuracy_test2
```
the accuracy in test data is 94%.
```{r}
# final model result
performance.poly <- data.frame(
  method = 'Polynomial Logistic',
  train_accuracy = accuracy_train2,
  test_accuracy = accuracy_test2
  )
performance.poly
```

# SVM
```{r}
#radial kernel
library(e1071)
set.seed(14)
tune.out <- tune(svm,price~.,data = baked_train1,
                 kernel="radial",
                 ranges=list(
                   cost=c(0.1,1,5,10,15,50,100),
                   gamma=c(0.01,0.05,0.1,0.5,1,2)))
summary(tune.out)
```
The best gamma is 0.01 and best cost is 5.

```{r}
fitsvm_radial<-svm(price~., 
                   data = baked_train1, 
                   kernel = "radial", 
                   cost = 5, 
                   gamma = 0.01)
```


```{r}
set.seed(14)
tune.out.poly<-tune(svm,price~.,data=baked_train1,
                    kernel="polynomial",
                    ranges=list(
                      cost=c(0.1,1,10,50,100,150,180),
                      degree=c(1,2,3,4,5)))
summary(tune.out.poly)
```
The best degree is 1 and best cost is 50.

```{r}
fitsvm_poly<-svm(price~., 
                 data = baked_train1, 
                 kernel = "polynomial", 
                 cost = 150, 
                 degree = 1)
```

```{r}
#svm training accuracy
predsvmradial_train<-predict(fitsvm_radial, newdata = baked_train1 )
table(true=baked_train1$price,pred=predsvmradial_train)
```
Radial Kernel Train Accuracy=(796+792)/1600=0.9925
```{r}
(svm_radial_train_accuracy <- sum(predsvmradial_train==baked_train1$price)/length(predsvmradial_train))
```

```{r}
#svm testing accuracy
predsvmradial_test<-predict(fitsvm_radial,newdata = baked_test1)
table(true=baked_test1$price,pred=predsvmradial_test)
```
Radial Kernel Test Accuracy=(198+191)/400=0.9725
```{r}
(svm_radial_test_accuracy <- sum(predsvmradial_test==baked_test1$price)/length(predsvmradial_test))
```

```{r}
#svm_poly training accuracy
predsvmpoly_train<-predict(fitsvm_poly, newdata = baked_train1 )
table(true=baked_train1$price,pred=predsvmpoly_train)
```
Polynomial Kernel Train Accuracy=(797+796)/1600=0.995625
```{r}
(svm_poly_train_accuracy <- sum(predsvmpoly_train==baked_train1$price)/length(predsvmpoly_train))
```

```{r}
#svm_poly testing accuracy
predsvmpoly_test<-predict(fitsvm_poly,newdata = baked_test1)
table(true=baked_test1$price,pred=predsvmpoly_test)
```
Polynomial Kernel Test Accuracy=(196+194)/400=0.975
```{r}
(svm_poly_test_accuracy <- sum(predsvmpoly_test==baked_test1$price)/length(predsvmpoly_test))
```


```{r}
# final model result
performance.svm_r <- data.frame(
  method = 'SVM Radial Kernel',
  train_accuracy = svm_radial_train_accuracy,
  test_accuracy = svm_radial_test_accuracy
  )
performance.svm_r
```

```{r}
# final model result
performance.svm_p <- data.frame(
  method = 'SVM Polynomial Kernel',
  train_accuracy = svm_poly_train_accuracy,
  test_accuracy = svm_poly_test_accuracy
  )
performance.svm_p
```

# Tree Methods
```{r}
library(rpart)       # direct engine for decision tree application
library(rpart.plot)  # for plotting decision trees
```

```{r}
dt1 <- rpart(
  formula=price~.,
  data= baked_train1,
  method="class"
)
rpart.plot(dt1)
```
```{r}
## control max depth
dt2<- rpart(
  formula = price ~ .,
  data    = baked_train1,
  method  = "class",
  control = list(maxdepth = 2)
)
rpart.plot(dt2)
```
```{r}
## control min node size
dt3 <- rpart(
  formula = price ~ .,
  data    = baked_train1,
  method  = "class",
  control = list(minbucket = round(0.1*nrow(baked_train1))) # at least 10%
)
rpart.plot(dt3)
```
```{r}
## try to grow the whole tree
dt4 <- rpart(
  formula = price ~ .,
  data    = baked_train1,
  method  = "class",
  control = list(minbucket=1) 
)
rpart.plot(dt4)
```

```{r}
dt5 <- rpart(
  formula = price ~ .,
  data    = baked_train1,
  method  = "class",
  control = list(minbucket=1,cp=0) 
)
rpart.plot(dt5)
```

```{r}
## get max depth of rpart tree
nodes <- as.numeric(rownames(dt5$frame))
max(rpart:::tree.depth(nodes))
```

# Random Forest
```{r}
library(h2o) # for resampling and model training
# h2o set-up
h2o.no_progress() # turn off h2o progress bars
h2o.init() # launch h2o
```

```{r}
library(ranger)
# number of features
n_features <- length(setdiff(names(baked_train1), "price"))
# train a default random forest model
rf1 <- ranger(
price ~ .,
data = baked_train1,
mtry = NULL,
splitrule = "gini",
seed = 14
)
# get OOB RMSE
(default_rmse <- sqrt(rf1$prediction.error))
```

```{r}
# create hyperparameter grid
hyper_grid <- expand.grid(
mtry = floor(n_features * c(.05, .15, .25, .333, .4)),
min.node.size = c(1, 3, 5, 10, 20),
replace = c(TRUE, FALSE),
sample.fraction = c(.5, .63, .8),
rmse = NA
)
```

```{r}
# execute full cartesian grid search
for(i in seq_len(nrow(hyper_grid))) {
# fit model for ith hyperparameter combination
fit <- ranger(
formula = price ~ .,
data = baked_train1,
num.trees = n_features * 10,
mtry = hyper_grid$mtry[i],
min.node.size = hyper_grid$min.node.size[i],
replace = hyper_grid$replace[i],
sample.fraction = hyper_grid$sample.fraction[i],
verbose = FALSE,
splitrule = "gini",
seed = 14
)
# export OOB error
hyper_grid$rmse[i] <- sqrt(fit$prediction.error)
}
```

The best mtry is 8 and min.node.size is 3.
```{r}
# assess top 10 models
hyper_grid %>%
arrange(rmse) %>%
mutate(perc_gain = (default_rmse - rmse) / default_rmse * 100) %>%
head(10)
```

```{r}
#h2o.no_progress()
#h2o.init(max_mem_size = "5g")
```

```{r}
# convert training data to h2o object
train_h2o <- as.h2o(baked_train1)
# set the response column to price
response <- "price"
# set the predictor names
predictors <- setdiff(colnames(baked_train1), response)
```

```{r}
h2o_rf1 <- h2o.randomForest(
x = predictors,
y = response,
training_frame = train_h2o,
ntrees = n_features * 10,
seed = 14
)
h2o_rf1
```
```{r}
# hyperparameter grid
hyper_grid <- list(
mtries = floor(n_features * c(.05, .15, .25, .333, .4)),
min_rows = c(1, 3, 5, 10),
max_depth = c(10, 20, 30),
sample_rate = c(.55, .632, .70, .80)
)
# random grid search strategy
search_criteria <- list(
strategy = "RandomDiscrete",
stopping_metric = "mse",
stopping_tolerance = 0.001, # stop if improvement is < 0.1%
stopping_rounds = 10 # over the last 10 models
)
```



```{r}
# perform grid search
random_grid <- h2o.grid(
algorithm = "randomForest",
grid_id = "rf_random_grid",
x = predictors,
y = response,
training_frame = train_h2o,
hyper_params = hyper_grid,
ntrees = n_features * 10,
seed = 14,
stopping_metric = "RMSE",
stopping_rounds = 10, # stop if last 10 trees added
stopping_tolerance = 0.005,# don’t improve RMSE by 0.5%
search_criteria = search_criteria
)
```

The best max_depth is 20, min.rows is 1 , and mtry is 8.
```{r}
random_grid_perf <- h2o.getGrid(
grid_id = "rf_random_grid",
sort_by = "mse",
decreasing = FALSE
)
random_grid_perf
```
```{r}
# re-run model with impurity-based variable importance
rf_impurity <- ranger(
formula = price ~ .,
data = baked_train1,
num.trees = 20,
mtry = 8,
min.node.size = 1,
sample.fraction = .80,
replace = FALSE,
importance = "impurity",
verbose = FALSE,
splitrule = "gini",
seed = 14
)
# re-run model with permutation-based variable importance
rf_permutation <- ranger(
formula = price ~ .,
data = baked_train1,
num.trees = 20,
mtry = 8,
min.node.size = 1,
sample.fraction = .80,
replace = FALSE,
importance = "permutation",
verbose = FALSE,
splitrule = "gini",
seed = 14
)
```

```{r}
library(vip)
p1 <- vip::vip(rf_impurity, num_features = 20, bar = FALSE)
p2 <- vip::vip(rf_permutation, num_features = 20, bar = FALSE)
gridExtra::grid.arrange(p1, p2, nrow = 1)
```

```{r}
predrfimpurity_train<-predict(rf_impurity, baked_train1 )
table(true=baked_train1$price,pred=predrfimpurity_train$predictions)
```
Random Forest impurity Train Accuracy=100%
```{r}
rf_train_accurarcy<- sum(baked_train1$price==predrfimpurity_train$predictions)/length(baked_train1$price)
rf_train_accurarcy
```


```{r}
predrfimpurity_test<-predict(rf_impurity, baked_test1 )
table(true=baked_test1$price,pred=predrfimpurity_test$predictions)
```
Random Forest impurity Train Accuracy=(194+191)/400=96.25%
```{r}
rf_test_accurarcy<- sum(baked_test1$price==predrfimpurity_test$predictions)/length(baked_test1$price)
rf_test_accurarcy
```

```{r}
predrfpermutation_train<-predict(rf_permutation, data = baked_train1,type="response")
table(true=baked_train1$price,pred=predrfpermutation_train$predictions)
```
Random Forest permutation Train Accuracy=100%

```{r}
predrfpermutation_test<-predict(rf_permutation, baked_test1 )
table(true=baked_test1$price,pred=predrfpermutation_test$predictions)
```
```{r}
# final model result
performance.rf <- data.frame(
  method = 'Random Forest',
  train_accuracy = rf_train_accurarcy,
  test_accuracy = rf_test_accurarcy
  )
performance.rf
```
```{r}
#h2o.shutdown(prompt = FALSE)
```

# GMB 
```{r include=FALSE}
#library(h2o)
#h2o.init(nthreads = -1) 
```

```{r}
# column names of x and y 
y <- "price"

# Identify the predictor columns
x <- setdiff(names(baked_train1), y)
```


```{r}
# Convert response to factor
baked_train1[,y] <- as.factor(baked_train1[,y])
#test[,y] <- as.factor(test[,y])

# convert train and test data frames to h2o objects
train.h2o <- as.h2o(baked_train1)
```
```{r}
# test data 
# Convert response to factor
baked_test1[,y] <- as.factor(baked_test1[,y])
#test[,y] <- as.factor(test[,y])

# convert train and test data frames to h2o objects
test.h2o <- as.h2o(baked_test1)
```

```{r eval=FALSE, include=FALSE}
fit.gbm <- h2o.gbm(
  x = x,
  y = y,
  training_frame = train.h2o,
  ntrees = 500,
  learn_rate = 0.1,
  nfolds = 10,
  max_depth = 5,
  sample_rate = 1.0,
  col_sample_rate_per_tree = 0.5,
  distribution = "bernoulli",
  stopping_rounds = 50,
  score_each_iteration = TRUE,
  stopping_metric ="misclassification",
  stopping_tolerance = 0,
  seed = 14
  )
```

```{r eval=FALSE, include=FALSE}
show(fit.gbm)
```

```{r eval=FALSE, include=FALSE}
gbm_s <- summary(fit.gbm)
```

# step1 tune learning rate
```{r}
# defined hyperparameter grid
hyper_grid_lr <- list(
  learn_rate = c(0.3,0.1, 0.05, 0.01, 0.005)             
)
```

```{r}
# perform grid search 
grid <- h2o.grid(
  algorithm = "gbm",
  grid_id = "gbm_lr_grid",
  x = x, 
  y = y,
  training_frame = train.h2o,
  hyper_params = hyper_grid_lr,
  ntrees = 500,
  max_depth = 5,
  min_rows = 5,
  nfolds = 10,
  distribution = "bernoulli",
  stopping_rounds = 50,
  score_each_iteration = TRUE,
  stopping_metric ="misclassification",
  stopping_tolerance = 0,
  seed = 14
)
```

```{r}
# collect the results and sort by our model performance metric of choice
lr_grid_perf <- h2o.getGrid(
  grid_id = "gbm_lr_grid", 
  sort_by = "accuracy", 
  decreasing = TRUE
)
```

```{r}
lr_grid_perf
```


```{r}
# get best learning rate
best_learning_rate <- lr_grid_perf@summary_table[1,1]
tune_shrinkage_result <- lr_grid_perf@summary_table[1,3]
best_learning_rate
```

# check how many trees does learning rate = 0.1 need 
```{r}
fit.gbm <- h2o.gbm(
  x = x,
  y = y,
  training_frame = train.h2o,
  ntrees = 500,
  learn_rate = best_learning_rate,
  nfolds = 10,
  max_depth = 5,
  sample_rate = 1.0,
  col_sample_rate_per_tree = 0.5,
  distribution = "bernoulli",
  stopping_rounds = 50,
  score_each_iteration = TRUE,
  stopping_metric ="misclassification",
  stopping_tolerance = 0,
  seed = 14
  )
```
```{r}
show(fit.gbm)
```
197 trees 

# step2 using best learning rate to tune tree-based methods hyperparameters
```{r}
# defined hyperparameter grid
hyper_grid_tree <- list(
  ntrees = c(50,100,300),
  max_depth = c(5,15,20),
  min_rows = c(5,15,20)
)
```

```{r}
# random grid search strategy
search_criteria <- list(
  strategy = "RandomDiscrete",
  stopping_metric = "misclassification",
  stopping_rounds = 20,         
  max_runtime_secs = 60*60      
)
```

```{r}
# perform grid search 
grid2 <- h2o.grid(
  algorithm = "gbm",
  grid_id = "gbm_tree_grid",
  x = x, 
  y = y,
  training_frame = train.h2o,
  learn_rate = best_learning_rate,  ### use best learning rate
  hyper_params = hyper_grid_tree,
  nfolds = 10,
  distribution = "bernoulli",
  search_criteria = search_criteria,
  seed = 14
)
```

```{r}
# collect the results and sort by our model performance metric of choice
tree_grid_perf <- h2o.getGrid(
  grid_id = "gbm_tree_grid", 
  sort_by = "accuracy", 
  decreasing = TRUE
)
```

```{r}
tree_grid_perf@summary_table
```
```{r}
# get best parameters
best_max_depth <- tree_grid_perf@summary_table[1,1]
best_min_rows <- tree_grid_perf@summary_table[1,2]
best_ntress <- tree_grid_perf@summary_table[1,3]
tune_tree_result <- tree_grid_perf@summary_table[1,5]
```

# step3 use basic gbm hyperparameters tune stochastic hyperparameters
```{r}
# refined hyperparameter grid
hyper_grid_stochastic <- list(
  sample_rate = c(0.5, 0.75, 1),              # row subsampling
  col_sample_rate = c(0.5, 0.75, 1),          # col subsampling for each split
  col_sample_rate_per_tree = c(0.5, 0.75, 1)  # col subsampling for each tree
)
```


```{r}
# perform grid search 
grid3 <- h2o.grid(
  algorithm = "gbm",
  grid_id = "stochastic_grid",
  x = x, 
  y = y,
  training_frame = train.h2o,
  hyper_params = hyper_grid_stochastic,
  ntrees = best_ntress,
  learn_rate = best_learning_rate,
  max_depth = best_max_depth,
  min_rows = best_min_rows,
  nfolds = 10,
  distribution = "bernoulli",
  search_criteria = search_criteria,
  seed = 14
)
```

```{r}
# collect the results and sort by our model performance metric of choice
stochastic_grid_perf <- h2o.getGrid(
  grid_id = "stochastic_grid", 
  sort_by = "accuracy", 
  decreasing = TRUE
)
```

```{r}
stochastic_grid_perf@summary_table
```


```{r}
# get best parameters
best_col_sample_rate <- stochastic_grid_perf@summary_table[1,1]
best_col_sample_rate_per_tree <- stochastic_grid_perf@summary_table[1,2]
best_sample_rate <- stochastic_grid_perf@summary_table[1,3]
tune_stochastic_result <- stochastic_grid_perf@summary_table[1,5]
```

# summarize tune result
```{r}
gbm_tune_result <- data.frame(
  tuned_hyperparamters = c('Shrinkage','Tree-based','Stochastic'),
  best_cv_accuracy = c(tune_shrinkage_result,tune_tree_result,tune_stochastic_result)
  )
gbm_tune_result$tuned_hyperparamters <- factor(gbm_tune_result$tuned_hyperparamters,levels = c('Shrinkage','Tree-based','Stochastic'))
gbm_tune_result
```


```{r}
ggplot(data = gbm_tune_result, aes(x = tuned_hyperparamters, y = best_cv_accuracy)) +
         #geom_bar(stat="identity") +
         geom_col(width = 0.5) + 
         geom_text(aes(label = best_cv_accuracy), vjust = 0) +
         coord_cartesian(ylim=c(0.95,1))
```



```{r}
# final model using best parameters
fit.gbm <- h2o.gbm(
  x = x,
  y = y,
  training_frame = train.h2o,
  validation_frame = test.h2o,
  learn_rate = best_learning_rate,
  max_depth = best_max_depth,
  min_rows = best_min_rows,
  ntrees = best_ntress,
  sample_rate = best_sample_rate,
  col_sample_rate_per_tree = best_col_sample_rate_per_tree,
  col_sample_rate = best_col_sample_rate,
  distribution = "bernoulli",
  stopping_rounds = 20,
  score_each_iteration = TRUE,
  stopping_metric ="misclassification",
  stopping_tolerance = 0,
  seed = 14
  )
```

# train data performance
```{r}
gbm_train_performance <-h2o.confusionMatrix(fit.gbm,metric="accuracy")
gbm_train_performance
```


```{r}
gbm_train_accuracy = 1- gbm_train_performance[3,3]
gbm_train_accuracy
```

# test data performance 
```{r}
gbm_test_performance <- h2o.confusionMatrix(fit.gbm,valid=TRUE,metric ="accuracy")
gbm_test_performance
```
```{r}
gbm_test_accuracy = 1- gbm_test_performance[3,3]
gbm_test_accuracy
```

```{r}
# final model performance
performance.gbm <- data.frame(
  method = 'GBM',
  train_accuracy = gbm_train_accuracy,
  test_accuracy = gbm_test_accuracy 
  )
```

```{r}
performance.gbm
```

```{r}
# Create custom predict function that returns the predicted values as a vector
pred <- function(object, newdata)  {
  results <- as.vector(h2o.predict(object, as.h2o(newdata)))
  return(results)
}
```


```{r}
library(vip)
vip(
  fit.gbm,
  train = as.data.frame(train.h2o),
  method = "permute",
  target = "price",
  metric = "accuracy",
  pred_wrapper = pred,
  scale = TRUE 
)
```

```{r}
# Custom prediction function wrapper
pdp_pred <- function(object, newdata)  {
  results <- mean(as.vector(h2o.predict(object,as.h2o(newdata))$p1))
  return(results)
}
```

```{r include=FALSE}
# Compute partial dependence values
pd_values_ram <- pdp::partial(
  fit.gbm,
  train = as.data.frame(train.h2o), 
  pred.var = "ram",
  pred.fun = pdp_pred,
  grid.resolution = 20
)
```
```{r include=FALSE}
# Compute partial dependence values
pd_values_bat <- pdp::partial(
  fit.gbm,
  train = as.data.frame(train.h2o), 
  pred.var = "battery_power",
  pred.fun = pdp_pred,
  grid.resolution = 20
)
```

```{r}
grid.arrange(
  autoplot(pd_values_ram, rug = TRUE, train = as.data.frame(train.h2o)),
  autoplot(pd_values_bat, rug = TRUE, train = as.data.frame(train.h2o)),
  ncol =2
)
```


# xgboost

```{r}
library(xgboost)
library(Matrix)
```

```{r}
y <- "price"
train.mx <- sparse.model.matrix(price ~ ., baked_train1)[,-1] 
test.mx <- sparse.model.matrix(price ~ ., baked_test1)[,-1]
dtrain <- xgb.DMatrix(train.mx, label = as.integer(baked_train1[,y])-1)
dtest <- xgb.DMatrix(test.mx, label = as.integer(baked_test1[,y])-1)
```


```{r}
#start_time = Sys.time()
#start_time 
set.seed(14)
system.time(mobile.xgb <- xgb.cv(
  data = dtrain,
  #label = getinfo(dtrain, 'label'),
  metrics = c('error'),
  nrounds = 300,
  objective = "binary:logistic",
  early_stopping_rounds = 50, 
  nfold = 10,
  params = list(
    eta = 0.1,
    max_depth = 5,
    min_child_weight = 15,
    subsample = 0.75,
    colsample_bytree = 1),
  verbose = 0
))

# maximize test CV accuracy
(default_accuracy_xgb <- 1-min(mobile.xgb$evaluation_log$test_error_mean))
```


```{r}
mobile.xgb$best_iteration
```

1. learning rate
```{r}
hyper_grid_eta <- expand.grid(
  eta = c(0.3,0.1, 0.05, 0.005), # typical 0.001-0.3
  accuracy = 0,          # a place to dump accuracy results
  trees = 0,          # a place to dump required number of trees
  time = 0
)
```

```{r}
# grid search  
start_time = Sys.time()
start_time 

for(i in seq_len(nrow(hyper_grid_eta))) {
  set.seed(14)
  train_time <- system.time({m <- xgb.cv(
    data = dtrain,
    #label = Y,
    nrounds = 1000,
    objective = "binary:logistic",
    early_stopping_rounds = 50, 
    metrics = c('error'),
    nfold = 10,
    verbose = 0,
    eta = hyper_grid_eta$eta[i]
  )})
  
  hyper_grid_eta$accuracy[i] <- 1- min(m$evaluation_log$test_error_mean)
  hyper_grid_eta$trees[i] <- m$best_iteration
  hyper_grid_eta$time[i]  <- train_time[["elapsed"]]
}
```
```{r}
#assess models
hyper_grid_eta %>%
  arrange(desc(accuracy)) %>%
  mutate(perc_gain = (accuracy-default_accuracy_xgb)/default_accuracy_xgb*100)
```

```{r}
#best performance
best_accuracy_eta <- max(hyper_grid_eta$accuracy)
best_accuracy_eta
```

2. tune tree-specific hyperparameters
```{r}
hyper_grid_tree_xgb <- expand.grid(
  eta = 0.3,
  max_depth = c(3,5,6,8,15), # typical 3-8
  min_child_weight = c(1,5,10,15), # typical 5-15
  accuracy = 0,          # a place to dump accuracy results
  trees = 0,          # a place to dump required number of trees
  time = 0
)
```

```{r}
# grid search  # total 252 combinations # 4 min
start_time = Sys.time()
start_time 

for(i in seq_len(nrow(hyper_grid_tree_xgb))) {
  set.seed(14)
  train_time <- system.time({m <- xgb.cv(
    data = dtrain,
    #label = Y,
    nrounds = 300,
    objective = "binary:logistic",
    early_stopping_rounds = 50, 
    metrics = c('error'),
    nfold = 10,
    verbose = 0,
    params = list(
      eta = hyper_grid_tree_xgb$eta[i], 
      max_depth = hyper_grid_tree_xgb$max_depth[i],
      min_child_weight = hyper_grid_tree_xgb$min_child_weight[i]
    ) 
  )})
  
  hyper_grid_tree_xgb$accuracy[i] <- 1- min(m$evaluation_log$test_error_mean)
  hyper_grid_tree_xgb$trees[i] <- m$best_iteration
  hyper_grid_tree_xgb$time[i]  <- train_time[["elapsed"]]
}
```

```{r}
#assess models
hyper_grid_tree_xgb %>%
  arrange(desc(accuracy)) %>%
  mutate(perc_gain = (accuracy-default_accuracy_xgb)/default_accuracy_xgb*100)
```
it seems that when min_child_weight = 1, the accuracy is the highest, not very sensitive to max_depth

```{r}
#best performance
best_accuracy_tree <- max(hyper_grid_tree_xgb$accuracy)
best_accuracy_tree
```

3. Explore stochastic GBM attributes.
```{r}
hyper_grid_stoch_xgb <- expand.grid(
  eta = 0.3,
  max_depth = 5, 
  min_child_weight = 1,
  subsample = c(0.5,0.75,1),
  colsample_bytree = c(0.5,0.75,1),
  accuracy = 0,          # a place to dump accuracy results
  trees = 0,          # a place to dump required number of trees
  time = 0
)
```

```{r}
# grid search  # total 252 combinations #  4 min
start_time = Sys.time()
start_time 

for(i in seq_len(nrow(hyper_grid_stoch_xgb))) {
  set.seed(14)
  train_time <- system.time({m <- xgb.cv(
    data = dtrain,
    #label = Y,
    nrounds = 300,
    objective = "binary:logistic",
    early_stopping_rounds = 50, 
    metrics = c('error'),
    nfold = 10,
    verbose = 0,
    params = list(
      eta = hyper_grid_stoch_xgb$eta[i], 
      max_depth = hyper_grid_stoch_xgb$max_depth[i],
      min_child_weight = hyper_grid_stoch_xgb$min_child_weight[i],
      subsample = hyper_grid_stoch_xgb$subsample[i],
      colsample_bytree = hyper_grid_stoch_xgb$colsample_bytree[i]
    ) 
  )})
  
  hyper_grid_stoch_xgb$accuracy[i] <- 1- min(m$evaluation_log$test_error_mean)
  hyper_grid_stoch_xgb$trees[i] <- m$best_iteration
  hyper_grid_stoch_xgb$time[i]  <- train_time[["elapsed"]]
}
```

```{r}
#assess models
hyper_grid_stoch_xgb %>%
  arrange(desc(accuracy)) %>%
  mutate(perc_gain = (accuracy-default_accuracy_xgb)/default_accuracy_xgb*100)
```

```{r}
#best performance
best_accuracy_stoch <- max(hyper_grid_stoch_xgb$accuracy)
best_accuracy_stoch
```
final model 
```{r}
set.seed(14)
fit.xgb1 <- xgboost(
  data = dtrain,
  nrounds = 35,
  eta = 0.3,
  max_depth = 5, 
  min_child_weight = 1,
  #subsample=0.75,
  #colsample_bytree=1,
  objective = "binary:logistic",
  verbose = 0
)
```


```{r}
train_pred <- predict(fit.xgb1, dtrain)
train_class <- as.integer(train_pred > 0.5)
xgb_train_accuracy1 <- sum(train_class==getinfo(dtrain, 'label'))/length(train_class)
xgb_train_accuracy1
```

```{r}
test_pred <- predict(fit.xgb1, dtest)
test_class <- as.integer(test_pred > 0.5)
xgb_test_accuracy1 <- sum(test_class==getinfo(dtest, 'label'))/length(test_class)
xgb_test_accuracy1
```

Above the result without examing the three regularization parmaters : gamma, lambda, and alpha.
```{r}
# hyperparameter grid 
hyper_grid_xgb <- expand.grid(
  eta = 0.3,
  max_depth = 5, 
  min_child_weight = 1,
  #subsample=0.75,
  #colsample_bytree=1,
  gamma = c(0, 1e-2, 0.1, 1, 10, 100, 1000),
  lambda = c(0, 1e-2, 0.1, 1, 100, 1000),
  alpha = c(0, 1e-2, 0.1, 1, 100, 1000),
  accuracy = 0,          # a place to dump accuracy results
  trees = 0,          # a place to dump required number of trees
  time = 0
)
```


```{r}
# grid search  # total 252 combinations #  8 min
start_time = Sys.time()
start_time 

for(i in seq_len(nrow(hyper_grid_xgb))) {
  set.seed(14)
  train_time <- system.time({m <- xgb.cv(
    data = dtrain,
    #label = Y,
    nrounds = 300,
    objective = "binary:logistic",
    early_stopping_rounds = 50, 
    metrics = c('error'),
    nfold = 10,
    verbose = 0,
    params = list(
      eta = hyper_grid_xgb$eta[i], 
      max_depth = hyper_grid_xgb$max_depth[i],
      min_child_weight = hyper_grid_xgb$min_child_weight[i],
      gamma = hyper_grid_xgb$gamma[i], 
      lambda = hyper_grid_xgb$lambda[i], 
      alpha = hyper_grid_xgb$alpha[i]
    ) 
  )})
  
  hyper_grid_xgb$accuracy[i] <- 1- min(m$evaluation_log$test_error_mean)
  hyper_grid_xgb$trees[i] <- m$best_iteration
  hyper_grid_xgb$time[i]  <- train_time[["elapsed"]]
}
```



```{r}
#assess top 10 models
hyper_grid_xgb %>%
  arrange(desc(accuracy)) %>%
  mutate(perc_gain = (accuracy-default_accuracy_xgb)/default_accuracy_xgb*100) %>%
  head(20)
```

```{r}
#best performance
best_accuracy_regul <- max(hyper_grid_xgb$accuracy)
best_accuracy_regul
```

# summarize tune result
```{r}
xgb_tune_result <- data.frame(
  tuned_hyperparamters = c('Shrinkage','Tree-based','Stochastic','Regularization'),
  best_cv_accuracy = c(best_accuracy_eta,best_accuracy_tree,best_accuracy_stoch,best_accuracy_regul)
  )
xgb_tune_result$tuned_hyperparamters <- factor(xgb_tune_result$tuned_hyperparamters,levels = c('Shrinkage','Tree-based','Stochastic','Regularization'))
xgb_tune_result
```

```{r}
ggplot(data = xgb_tune_result, aes(x = tuned_hyperparamters, y = best_cv_accuracy)) +
         #geom_bar(stat="identity") +
         geom_col(width = 0.5) + 
         geom_text(aes(label = best_cv_accuracy), vjust = 0) +
         coord_cartesian(ylim=c(0.97,0.99))
```
```{r}
set.seed(14)
fit.xgb2 <- xgboost(
  data = dtrain,
  nrounds = 64,
  eta = 0.3,
  max_depth = 5, 
  min_child_weight = 1,
  gamma = 0,
  lambda =1, 
  alpha = 0.01,
  objective = "binary:logistic",
  verbose = 0
)
```

```{r}
train_pred <- predict(fit.xgb2, dtrain)
train_class <- as.integer(train_pred > 0.5)
xgb_train_accurarcy2 <- sum(train_class==getinfo(dtrain, 'label'))/length(train_class)
xgb_train_accurarcy2
```

```{r}
test_pred <- predict(fit.xgb2, dtest)
test_class <- as.integer(test_pred > 0.5)
xgb_test_accurarcy2 <- sum(test_class==getinfo(dtest, 'label'))/length(test_class)
xgb_test_accurarcy2
```


## four important predictors
```{r}
vip::vip(fit.xgb2,scale=TRUE)
```

## pdp plots
```{r}
# Custom prediction function wrapper
xgb_pred <- function(object, newdata)  {
  results <- mean(as.vector(predict(object, newdata)))
  return(results)
}
```

```{r}
# Compute partial dependence values
xgb_values_ram <- pdp::partial(
  fit.xgb2,
  train = train.mx,
  pred.var = "ram",
  pred.fun = xgb_pred,
  grid.resolution = 20
)
```

```{r}
# Compute partial dependence values
xgb_values_bat <- pdp::partial(
  fit.xgb2,
  train = train.mx,
  pred.var = "battery_power",
  pred.fun = xgb_pred,
  grid.resolution = 20
)
```


```{r}
vip::grid.arrange(
  autoplot(xgb_values_ram, rug = TRUE, train = train.mx),
  autoplot(xgb_values_bat, rug = TRUE, train = train.mx),
  ncol=2
)
```

```{r}
# final model result
performance.xgb <- data.frame(
  method = 'XGBoost',
  train_accuracy = xgb_train_accuracy1,
  test_accuracy = xgb_test_accuracy1
  )
performance.xgb
```


```{r}
comparasion <- rbind(performance.log,performance.poly,performance.rf,performance.svm_p,performance.svm_r,performance.gbm,performance.xgb)
comparasion$method <- factor(comparasion$method ,levels = c('Logistic','Polynomial Logistic','Random Forest','SVM Polynomial Kernel','SVM Radial Kernel','GBM','XGBoost'))
comparasion
```

```{r}
library(reshape2)
cp <- melt(comparasion, id.vars='method')
#cp$value <- scales::percent(cp$value, 0.01)
```

```{r}
ggplot(data = cp, aes(x = method, y = value,fill=variable)) +
         geom_bar(stat='identity', position='dodge')+
         #scale_fill_grey() +
         scale_fill_manual(values=c('grey','orange'))+ #'#DF8344'
         #,fill=c('grey','orange')
         # geom_text(aes(label = round(value,4)),vjust = 0.5) +
         theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
         coord_cartesian(ylim=c(0.93,1))
```

```{r}
# good practice to shut down h2o environment
h2o.shutdown(prompt = FALSE)
```